---
title: "Ruka INIKA data"
author: "Ruka"
date: "2024-11-18"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# course part 1 - Laboratory
- we selected 20 isolates
- 6 isolates are ***Klebsiella spp.*** isolates
- 6 isolates are *Escherichia coli* isolates
- 8 isolates are *Salmonella spp.* isolates

## Istallation of packages

```{r}
```

```{r}
install.packages("here")
install.packages("DiagrammeR")
```

## Load the package

ctrl + alt + i (shortcut to keyboard)

```{r}
angry_or_not <- "not"
```
gives the content of the object (container)

```{r}
angry_or_not
```

Madelaine, are you angry or not? `r print(angry_or not)`

```{r}

```

#working with your data

```{r}
here::here()
```

```{r}
path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```
```{r}
human_data <-
  readr::read_csv2(path_data)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
glimpse(human_data)
str(human_data)
```

```{r}
library(tidyverse)
original_names <- colnames(human_data) 
original_names
```

```{r}
new_names <- str_replace_all(original_names, " ", "_")
new_names
new_names[88]
new_names
```

regex pattern
```{r}
new_names2 <-str_replace_all(new_names, "[?/,;.*()-]","_")
new_names2[88]
```

```{r}
colnames(human_data) <- new_names2
colnames(human_data)[88]
```

-selecting columns
```{r}
human_data %>%
  select(INIKA_OH_TZ_ID,Age__yrs_, Gender) %>%
  head()
```

select(INIKA_OH_TZ_ID, Age__yrs, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you, 
         Who_is_your_caretaker, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker, 
         Have_you_ever_heard_about_AMR, If_yes__how_did_you_get_this_information, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time, 
         If_yes__where_did_you_get_these_drugs_from,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber,
         GPS_coordinates_latitude, GPS_coordinates_longitude) 

human_data_selection %>% View()
human_data_selection %>% head()

Viewing the data or better an abstract of your data. Notion of pipe and selecting columns

```{r}
glimpse(human_data_selection)
str(human_data_selection)
```

```{r}
human_data_selection %>%
  select(-specify_if_other_district) %>%
  mutate_at(vars(INIKA_OH_TZ_ID) , as.character) %>%
  str()
```

-looking at the content of single column
